---
title: "Project 1"
author: "Archit Singh, Sreeram Kondapalli"
date: "2023-10-17"
output:
    ioslides_presentation:
        widescreen: yes
        smaller: yes
        transition: default
---
<style type="text/css">
body p, div, h1, h2, h3, h4, h5 {
    color: black;
    font-family: Modern Computer Roman;
}
slides > slide.title-slide hgroup h1 {
    color: #8C1D40; <!--the maroon color-->
    }
h2 {
    color: #8C1D40; <!-- the maroon color-->
}
slides > slide {
    background: linear-gradient(#ffffff, #ffffff 85%, #ffffff);
    background-color: white;
}
.left-slide-content {
  width: 50%;
  float: left;
}

.right-slide-content {
  width: 50%;
  float: right;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
```

## Introduction

- The purpose of this project is to explore how the rise of AI tools like ChatGPT have impacted the education industry and the
technology industry as a whole. We had the following guiding questions throughout this project.
  - Do the students in the survey believe that AI will have a large impact on education.?
  - How was Chegg's stock doing before the release of ChatGPT?
  - How quickly did ChatGPT reach 1 million users?
  - How does Chegg's stock price compare to ChatGPT's user growth?
  - What effect does the adoption of AI have on the stock price of publicly traded companies?

## Methods and Datasets

### Students' Perspective on AI

This dataset surveyed 91 students on their feelings about AI. There were many questions asked in this survey such as 
how the students learn about AI and when they think of AI what do they feel. However the main questions we are focusing on that were
asked in this survey, relate to our original question about whether the students believe that AI will have a large impact on education, 
these were the following relevant questions that help answer our questions when analyzed:

- Question 6: In which areas do you think AI would have a big impact? With their options being: Education, Medicine, Agriculture, Constructions, 
Marketing,Public administration, Art
- Question 7: On a scale of 1 to 10, how useful do you think AI would be in the educational process? (1- not useful at all, 10-extremely useful)

## Question 6: In which areas do you think AI would have a big impact? 

```{r}
students_survey_ai <- read.csv("Survey_AI.csv")
# Bar plot for Q6
# Separate the responses to question 6
df_q6 <- students_survey_ai %>%
  select(Q6.Domains) %>%
  separate_rows(Q6.Domains, sep = ";") %>%
  mutate(Q6.Domains = str_to_lower(str_trim(Q6.Domains)))

# Count the frequency of each domain
df_q6_count <- df_q6 %>%
  group_by(Q6.Domains) %>%
  summarise(Count = n())

# Plot the data
ggplot(df_q6_count, aes(x = reorder(Q6.Domains, Count), y = Count)) +
  geom_bar(stat = "identity", fill = "steelblue", width = 0.7) +
  coord_flip() +
  theme_minimal(base_size = 15) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 1, vjust = 0.5, color = "black"),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)),
    axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
    plot.title = element_text(hjust = 0.5),
    plot.caption = element_text(hjust = 1, size = 8)
  ) +
  labs(
    x = "Count",
    y = "Industries",
    title = "Impact Of AI Across Different Industries",
    caption = "Source: Student Survey on AI in Education"
  )
```

## Question 7: On a scale of 1 to 10, how useful do you think AI would be in the educational process?

```{r}
# Bar plot for Q7
ggplot(students_survey_ai, aes(x = Q7.Utility_grade)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "black") +
  labs(x = "Rating (1-10)", y = "Count", title = "How Useful Do You Think AI Would Be In The\nEducational Process")+
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 14, face = "bold"),
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.caption = element_text(hjust = 1, size = 8)
  )
```

## Chegg Stock

```{r}
# Load in data
chegg_stock_price <- read.csv("CHGG.csv")

# Ensure the Date column is a Date type
chegg_stock_price$Date <- as.Date(chegg_stock_price$Date, format="%Y-%m-%d")

# Filter the data
filtered_chegg_stock_price <- chegg_stock_price %>%
  filter(Date >= as.Date("2022-05-01") & Date <= as.Date("2022-11-01"))

# Enhanced graph
ggplot(data = filtered_chegg_stock_price, aes(x = Date, y = Close)) +
  geom_line(color = "blue", size = 1.5) +
  geom_point(color = "red", size = 3) +
  labs(title = "Chegg Stock Price 6 months before the release of ChatGPT",
       x = "Month",
       y = "Closing Price ($)",
       caption = "Source: Yahoo Finance") +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "gray90", size = 0.5),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none",
    plot.caption = element_text(hjust = 1, size = 8)
  ) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y")
```

## ChatGPT User Growth

<div class="left-slide-content">

```{r, out.width='100%'}
# Read data from the CSV file
chatgpt_users <- read.csv("Number of ChatGPT users recorded over past months.csv")
#Renaming for simplicity
chatgpt_users <- chatgpt_users %>%
  rename(TotalVisits = 'Number.of.ChatGPT.users.recorded.over.past.months')
chatgpt_users$Date <- as.Date(chatgpt_users$Date, format="%Y-%m-%d")

label_format_millions <- function(x) {
  scales::number_format(scale = 1e-6, suffix = "m")(x)
}

ggplot(chatgpt_users, aes(x=Date, y=TotalVisits)) +
  geom_point(color = "red", size = 3, shape = 19, alpha = 0.7) +
  geom_line(color = "blue", size = 1.2) +
  labs(title="Number of ChatGPT Users Over Time",
       subtitle="Monthly data from Nov 2022 to Aug 2023",
       x="Date", y="Total Visits (in millions)",
       caption = "Source: SimilarWeb") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  scale_y_continuous(labels = label_format_millions) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_text(size = 14, face = "bold"),
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank()
  )
```
</div>

<div class="right-slide-content">
```{r, out.width='100%'}
one_mil_users <- read.csv("time-taken-by-services-to-reach-1M.csv")
ggplot(one_mil_users, aes(x = reorder(Service, Days_to_reach_1M_users), y = Days_to_reach_1M_users)) +
  geom_bar(stat = "identity", fill = "#3498db", width = 0.7) +
  geom_text(aes(label = Days_to_reach_1M_users), position = position_dodge(0.5), hjust = -0.1, vjust = 0.5, color = "black", lineWidth = 4) +
  labs(
    title = "Days Taken by Services to Reach 1 Million Users",
    x = "Service",
    y = "Days to reach 1M users",
    caption = "Source: Nerdy Nav"
  ) +
  coord_flip() +
  ylim(0, max(one_mil_users$Days_to_reach_1M_users) + 100) +
  theme_light(base_size = 14) +
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    plot.title = element_text(hjust = 1, size = 14, face = "bold")
  )
```
</div>

## Chegg Stock VS ChatGPT User Growth

```{r, out.width='100%'}
chatgpt_users$Date <- as.Date(chatgpt_users$Date)
filtered_chegg_stock_price <- chegg_stock_price %>%
  filter(Date >= as.Date("2022-05-01") & Date <= as.Date("2023-08-01"))
label_format_millions <- function(x) {
  scales::number_format(scale = 1e-6, suffix = "m")(x)
}

# Set up a scale factor based on the max values of both datasets
scale_factor <- max(chatgpt_users$TotalVisits) / max(filtered_chegg_stock_price$Close)

# Begin the ggplot visualization
p <- ggplot() +

  # Plot ChatGPT User Growth with thicker line and color adjustments
  geom_line(data = chatgpt_users, 
            aes(x = Date, y = TotalVisits, color = "ChatGPT User Growth"), 
            size = 1.2, alpha = 0.8) +
  
  # Plot Chegg Stock Price with adjusted scale factor, thicker line, and color adjustments
  geom_line(data = filtered_chegg_stock_price, 
            aes(x = Date, y = Close * scale_factor, color = "Chegg Stock Price"), 
            size = 1.2, alpha = 0.8) +

  # Set primary y-axis for ChatGPT User Growth with label format and secondary y-axis for Chegg Stock Price
  scale_y_continuous(name = "ChatGPT User Growth", 
                     labels = label_format_millions, 
                     sec.axis = sec_axis(~./scale_factor, name = "Chegg Stock Price")) +

  # Define graph title, subtitle, and x-axis label
  labs(title = "Chegg Stock Price vs ChatGPT User Growth", 
       subtitle = "Visual representation of growth trends over time", 
       x = "Date", 
       caption = "Data source: Yahoo Finance & SimilarWeb") +
  
  # Use a light grid theme for a professional look
  theme_light() +

  # Increase the title font size and style
  theme(plot.title = element_text(face = "bold", size = 16),
        plot.subtitle = element_text(face = "italic", size = 12),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.position = "bottom",
        legend.key.size = unit(1, "cm")) +

  # Manually set the colors for the two lines with a pleasant color palette
  scale_color_manual(values = c("Chegg Stock Price" = "#1f77b4", "ChatGPT User Growth" = "#ff7f0e"))

# Display the plot
print(p)
```

## Stock of NVDA and MSFT

<div class="left-slide-content">

```{r, out.width='100%'}
NVDA <- read.csv("NVDA.csv")

# Convert the 'Date' column to a Date object
NVDA$Date <- as.Date(NVDA$Date)

ggplot(NVDA, aes(x = Date, y = Close)) +
  geom_line(color = "green", size = 1) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  scale_y_continuous(labels = scales::dollar_format(scale = 1)) +
  labs(
    title = "NVIDIA Stock Price Over Time",
    x = "Date",
    y = "Close Price (USD)",
    caption = "Source: Stock Market Data"
  ) +
  theme_light(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    plot.caption = element_text(size = 8, hjust = 1)
  )
```
</div>

<div class="right-slide-content">

```{r, out.width='100%'}
MSFT <- read.csv("MSFT.csv")

# Convert the 'Date' column to a Date object
MSFT$Date <- as.Date(MSFT$Date)

ggplot(MSFT, aes(x = Date, y = Close)) +
  geom_line(color = "green", size = 1.2) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  labs(
    title = "Microsoft Stock Price Over Time",
    subtitle = "Based on Close Price",
    x = "Date",
    y = "Close Price ($)"
  ) +
  theme_light(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 14),
    axis.title = element_text(face = "bold", size = 14)
  ) +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y") 
```
</div>

## Stock of GOOGL

```{r}
GOOG <- read.csv("GOOGL.csv")

# Convert the 'Date' column to a Date object
GOOG$Date <- as.Date(GOOG$Date)

# Create a line plot using ggplot2 with a red line
ggplot(GOOG, aes(x = Date, y = Close)) +
  geom_line(color = "red", size = 1.2) + # Adjusted color and size for better visualization
  labs(
    title = "Google Stock Price Over Time",
    subtitle = "Based on Close Price", 
    x = "Date",
    y = "Close Price ($)"
  ) +
  theme_light(base_size = 12) + # Switched to theme_light for a cleaner look
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 14),
    axis.title = element_text(face = "bold", size = 14),
  )
```

## Results

- Surveyed Students believe that AI will have a major impact on education.
- Before the release of generative AI, Chegg's stock had never been higher meaning students were already
using services to help them in school.
- ChatGPT reached 1 million users in 5 days showing how they filled a gap that the market was looking for.
- Chegg's stock exactly started dropping since the release of ChatGPT
- All the companies that have released generative AI products have seen a major increase in their stocks.

The results of this project show that both socially and economically AI has already started impacting our lives

## Conclusion

- AI tools, like ChatGPT, have significantly influenced the education and technology sectors, offering a competitive advantage to early adopters.
- Companies such as Nvidia and Microsoft have shown improved market performance post their AI ventures, suggesting a model for others aiming for growth and improved services.
- Survey data indicates a positive impact of AI on educational processes, aligning with market trends.
- Traditional educational platforms like Chegg have seen a decline in stock value compared to the rise of AI-driven solutions like ChatGPT.
- The data emphasizes the importance of AI in the current digital landscape, suggesting a need for its broader integration across sectors for innovation and better service delivery.

## Further Study

- A cross-sector comparison is needed to understand AI's broader impact, especially in fields like healthcare, agriculture, and public administration.
- Ethical concerns related to AI, especially in education, need rigorous examination, including issues of data privacy, security, and biases.
- Policy and regulation dynamics concerning AI adoption present significant research opportunities, with potential effects on innovation.
- The economic effects of AI, including job displacement and broader market implications, warrant thorough investigation.
- These comprehensive studies will pave the way for informed decision-making in maximizing AI's benefits responsibly.
