---
title: "Project 1"
author: "Archit Singh, Sreeram Kondapalli"
output: 
  pdf_document: 
    toc: true
date: "2023-10-12"
---

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
```


# Introduction

The purpose of this project is to explore how the rise of AI tools like ChatGPT have impacted the education industry and the
technology industry as a whole. To do this the project will first explore how the education industry was faring before the release 
of AI LLMs (Large Language Models) like ChatGPT, by analyzing the stock price of a company focused on education, Chegg. Then, after 
its release we will see how ChatGPT's user growth affected Chegg's stock price. Data about many other companies stock as well as
many related topics about how students feel about AI will also be discussed as context for this study. Here are some basic questions
that will be answered in this project:

- Do the students in the survey believe that AI will have a large impact on education.?
- How was Chegg's stock doing before the release of ChatGPT?
- How quickly did ChatGPT reach 1 million users?
- How does Chegg's stock price compare to ChatGPT's user growth?
- What effect does the adoption of AI have on the stock price of publicly traded companies?

# Methods/ Datasets

In order to thoroughly analyze the topic, various datasets are required. The datasets range from stock market data of relevant companies
to user growth data for ChatGPT, as well as a student survey on how they feel about AI. The finance data for all of the companies' stocks
was gathered using the historical data feature of Yahoo Finance which allows you to download CSV files of a company's stock information in
a specified time frame. The data about ChatGPT's user growth and how long it took to reach 1 million users comes from an analytics firm, 
SimilarWeb, whose data we then compiled into a CSV file ourselves. The survey about how students feel about AI comes from Kaggle and it 
includes many questions that was asked to 90 students about various topics related to AI. If you are interested in learning more about any
of the datasets then they are all linked in the Appendix. Most of these datasets were already cleaned but there was some filtering and dplyr
verbs used in order to plot the results of the data.

## Students' Perspective on AI

We begin with a dataset that surveyed 91 students on their feelings about AI. There were many questions asked in this survey such as 
how the students learn about AI and when they think of AI what do they feel. However the main questions we are focusing on that were
asked in this survey, relate to our original question about whether the students believe that AI will have a large impact on education, 
these were the following relevant questions that help answer our questions when analyzed:

- Question 6: In which areas do you think AI would have a big impact? With their options being: Education, Medicine, Agriculture, Constructions, 
Marketing,Public administration, Art
- Question 7: On a scale of 1 to 10, how useful do you think AI would be in the educational process? (1- not useful at all, 10-extremely useful)

```{r}
students_survey_ai <- read.csv("Survey_AI.csv")
# Bar plot for Q6
# Separate the responses to question 6
df_q6 <- students_survey_ai %>%
  select(Q6.Domains) %>%
  separate_rows(Q6.Domains, sep = ";") %>%
  mutate(Q6.Domains = str_to_lower(str_trim(Q6.Domains)))

# Count the frequency of each domain
df_q6_count <- df_q6 %>%
  group_by(Q6.Domains) %>%
  summarise(Count = n())

# Plot the data
ggplot(df_q6_count, aes(x = reorder(Q6.Domains, Count), y = Count)) +
  geom_bar(stat = "identity", fill = "steelblue", width = 0.7) +
  coord_flip() +
  theme_minimal(base_size = 15) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 1, vjust = 0.5, color = "black"),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)),
    axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
    plot.title = element_text(hjust = 0.5),
    plot.caption = element_text(hjust = 1, size = 8)
  ) +
  labs(
    x = "Count",
    y = "Industries",
    title = "Impact Of AI Across Different Industries",
    caption = "Source: Student Survey on AI in Education"
  )
```

This bar plot shows us how the participants responded to the question "In which areas do you think AI would have a big impact?", we see that
medicine is the most closely followed by education. Around 60 students responded that they believe education will be impacted by AI, which is
about 2/3 of the survey participants. This partially answers our major question of whether the students believe that AI will have a large impact
on education, by indicating that most students do believe this is the case. To dig deeper into this question, we look at question 7 that 
specifically asked the participants about this. We wanted to investigate education further and how the participants felt about AI in education 
with question 7.

```{r}
# Bar plot for Q7
ggplot(students_survey_ai, aes(x = Q7.Utility_grade)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "black") +
  labs(x = "Rating (1-10)", y = "Count", title = "How Useful Do You Think AI Would Be In The\nEducational Process")+
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 14, face = "bold"),
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.caption = element_text(hjust = 1, size = 8)
  )
```

We can see from this bar plot that most participants believe that AI will have a large impact on education as 10 was the most frequent number.
It indicates that most students that participated in this survey believe that AI is going to be super helpful in the education sector. Around
20 students chose a value of 10 indicating how they feel that AI can impact their education.

## Chegg Stock

Chegg is a company that offers many services related to education such as renting textbooks and homework solutions. Chegg is only one of the homework 
help companies that is publicly traded, so it is a perfect candidate that can represent the older way that students would learn. A future study could analyze
some of the user loss of other homework help companies to see whether the same pattern held with them. Before the release of ChatGPT, we will see how Chegg's 
stock was on the rise. The ethics of this company and using their products aside before the release of AI tools, students turned to websites such as Chegg to 
help them in their school work. There were many external factors that might have contributed to the rise of Chegg's stock, such as the pandemic and students 
not being able to learn the material as well  We can see that Chegg stock was at it's peak before November of 2022.

```{r}
# Load in data
chegg_stock_price <- read.csv("CHGG.csv")

# Ensure the Date column is a Date type
chegg_stock_price$Date <- as.Date(chegg_stock_price$Date, format="%Y-%m-%d")

# Filter the data
filtered_chegg_stock_price <- chegg_stock_price %>%
  filter(Date >= as.Date("2022-05-01") & Date <= as.Date("2022-11-01"))

# Enhanced graph
ggplot(data = filtered_chegg_stock_price, aes(x = Date, y = Close)) +
  geom_line(color = "blue", size = 1.5) +
  geom_point(color = "red", size = 3) +
  labs(title = "Chegg Stock Price 6 months before the release of ChatGPT",
       x = "Month",
       y = "Closing Price ($)",
       caption = "Source: Yahoo Finance") +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "gray90", size = 0.5),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none",
    plot.caption = element_text(hjust = 1, size = 8)
  ) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y")
```

This line plot of Chegg's stock price over the months before the release of ChatGPT show that Chegg was doing better than ever before whether because of external
factors or just an exorbitant amount of students that needed help signed up for the service. Perhaps this graph can be extrapolated to the other companies that are
in the education industry to say that they were also all doing very well financially before the release of generative AI.

## ChatGPT User Growth 

Open AI is a research lab that is focused on AI. In November of 2022, coinciding with Chegg's best month for its stock, ChatGPT was released. ChatGPT, which stands for
Chat Generative Pre-trained Transformer, is a large language model that was trained on a really large dataset of the internet. It is what is known as generative AI 
and it is able to confidently try almost all tasks you ask it to do. Many students over the months since its release have adopted it to help them with their homework. 
Open AI is not a publicly traded company and because of this we cannot analyze its stock price. However, we can analyze its user growth over the months since its release 
by an analytical firm called SimilarWeb. The following is the user growth of ChatGPT as plotted over the months since its release.

```{r}
# Read data from the CSV file
chatgpt_users <- read.csv("Number of ChatGPT users recorded over past months.csv")
#Renaming for simplicity
chatgpt_users <- chatgpt_users %>%
  rename(TotalVisits = 'Number.of.ChatGPT.users.recorded.over.past.months')
chatgpt_users$Date <- as.Date(chatgpt_users$Date, format="%Y-%m-%d")

label_format_millions <- function(x) {
  scales::number_format(scale = 1e-6, suffix = "m")(x)
}

ggplot(chatgpt_users, aes(x=Date, y=TotalVisits)) +
  geom_point(color = "red", size = 3, shape = 19, alpha = 0.7) +
  geom_line(color = "blue", size = 1.2) +
  labs(title="Number of ChatGPT Users Over Time",
       subtitle="Monthly data from Nov 2022 to Aug 2023",
       x="Date", y="Total Visits (in millions)",
       caption = "Source: SimilarWeb") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  scale_y_continuous(labels = label_format_millions) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_text(size = 14, face = "bold"),
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank()
  )
```

ChatGPT's user growth almost seems exponential for the first couple months since it was released. In fact let us look at how long popular services took to reach 1 million 
users vs how long ChatGPT took. The release of an AI model such as ChatGPT must have shown the world and the established companies that specialized in answering people's
questions (Google), how powerful that generative AI can be and how they needed to develop their answer to this new technology, in order to stay relevant. In fact, ChatGPT's
growth was so momentous we think you need some context as to how big it really was and so we present the following data about how long popular services took to cross the 
threshold of a million users.

```{r}
one_mil_users <- read.csv("time-taken-by-services-to-reach-1M.csv")
ggplot(one_mil_users, aes(x = reorder(Service, Days_to_reach_1M_users), y = Days_to_reach_1M_users)) +
  geom_bar(stat = "identity", fill = "#3498db", width = 0.7) +
  geom_text(aes(label = Days_to_reach_1M_users), position = position_dodge(0.5), hjust = -0.1, vjust = 0.5, color = "black", lineWidth = 4) +
  labs(
    title = "Days Taken by Services to Reach 1 Million Users",
    x = "Service",
    y = "Days to reach 1M users",
    caption = "Source: Nerdy Nav"
  ) +
  coord_flip() +
  ylim(0, max(one_mil_users$Days_to_reach_1M_users) + 100) +
  theme_light(base_size = 14) +
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    plot.title = element_text(hjust = 1, size = 14, face = "bold")
  )
```

This bar plot shows us how quickly ChatGPT actually reached one million users, and the answer is 5 days. Compared to some other popular services today such as Netflix that
took over one thousand days. Some of the companies that have since come out with competitors to ChatGPT are also on this list such as Facebook/ Instagram (Meta). It is interesting
that even a company such as Meta could not beat Open AI to the punch for releasing generative AI to the public, even though they have maintained dominance for many years and had
one million users on both Facebook and Instagram for a very long time as well.
 
## ChatGPT User growth VS Chegg stock

Finally, to answer one of our original questions: "How does Chegg's stock price compare to ChatGPT's user growth?". We plot the following line plot to visually see the impact
that ChatGPT had on the stock price of Chegg. The result of the line plot will make clear how much of an impact ChatGPT's release had on the bottom line of Chegg and whether
it is a good indicator of what the students said in the survey earlier about ChatGPT having a major impact on education.

```{r}
chatgpt_users$Date <- as.Date(chatgpt_users$Date)
filtered_chegg_stock_price <- chegg_stock_price %>%
  filter(Date >= as.Date("2022-05-01") & Date <= as.Date("2023-08-01"))
label_format_millions <- function(x) {
  scales::number_format(scale = 1e-6, suffix = "m")(x)
}

# Set up a scale factor based on the max values of both datasets
scale_factor <- max(chatgpt_users$TotalVisits) / max(filtered_chegg_stock_price$Close)

# Begin the ggplot visualization
p <- ggplot() +

  # Plot ChatGPT User Growth with thicker line and color adjustments
  geom_line(data = chatgpt_users, 
            aes(x = Date, y = TotalVisits, color = "ChatGPT User Growth"), 
            size = 1.2, alpha = 0.8) +
  
  # Plot Chegg Stock Price with adjusted scale factor, thicker line, and color adjustments
  geom_line(data = filtered_chegg_stock_price, 
            aes(x = Date, y = Close * scale_factor, color = "Chegg Stock Price"), 
            size = 1.2, alpha = 0.8) +

  # Set primary y-axis for ChatGPT User Growth with label format and secondary y-axis for Chegg Stock Price
  scale_y_continuous(name = "ChatGPT User Growth", 
                     labels = label_format_millions, 
                     sec.axis = sec_axis(~./scale_factor, name = "Chegg Stock Price")) +

  # Define graph title, subtitle, and x-axis label
  labs(title = "Chegg Stock Price vs ChatGPT User Growth", 
       subtitle = "Visual representation of growth trends over time", 
       x = "Date", 
       caption = "Data source: Yahoo Finance & SimilarWeb") +
  
  # Use a light grid theme for a professional look
  theme_light() +

  # Increase the title font size and style
  theme(plot.title = element_text(face = "bold", size = 16),
        plot.subtitle = element_text(face = "italic", size = 12),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.position = "bottom",
        legend.key.size = unit(1, "cm")) +

  # Manually set the colors for the two lines with a pleasant color palette
  scale_color_manual(values = c("Chegg Stock Price" = "#1f77b4", "ChatGPT User Growth" = "#ff7f0e"))

# Display the plot
print(p)
```

This line plot basically answers our question about how Chegg's stock price coincides with the growth of ChatGPT. On November 30, 2022, the first version of ChatGPT was released and 
right after it was released we can see that Chegg's stock has been on the decline. In this plot, we can see that Chegg's stock basically coincides dropping exactly with when ChatGPT 
releases and it has been on the downturn ever since then. Whereas, in contrast to Chegg, ChatGPT has been growing until recently until it has started plateauing. There can be many factors
that can explain ChatGPT's user plateau/ decline, such as the release of GPT 4 as a subscription model, or the ethical concerns with AI. As well as a ban from some countries on ChatGPT.

## Stock of Companies That Released Public AI

On the topic of AI's growth we now change focus onto companies that have followed OpenAI's lead and released generative AI related products.These companies show how profitable AI has 
been for them and how it can impact many industries and not just education. The following are the companies that announced generative AI, beginning with Nvidia, Nvidia, makes graphics 
chips that have AI accelerator cards on board so they are great for running AI models. Nvidia has made a great fortune from this AI boom and by supplying many companies that want to release 
AI products with their chips .

```{r}
NVDA <- read.csv("NVDA.csv")

# Convert the 'Date' column to a Date object
NVDA$Date <- as.Date(NVDA$Date)

ggplot(NVDA, aes(x = Date, y = Close)) +
  geom_line(color = "green", size = 1) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  scale_y_continuous(labels = scales::dollar_format(scale = 1)) +
  labs(
    title = "NVIDIA Stock Price Over Time",
    x = "Date",
    y = "Close Price (USD)",
    caption = "Source: Stock Market Data"
  ) +
  theme_light(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    plot.caption = element_text(size = 8, hjust = 1)
  )
```

Nvidia's stock shows huge growth since the release of ChatGPT in November of 2022. Seems like Nvidia's stock price went from $200 to $450 over the period that we are analyzing.
Nvidia's market cap actually crossed one trillion dollars because of how popular their chips have been. Many people will only know Nvidia from their graphics cards made to run
games on the personal computers of people playing computer games however that is a very small business for Nvidia compared to how much money they make from their AI accelerator
cards. Next we analyze Microsoft:

```{r}
MSFT <- read.csv("MSFT.csv")

# Convert the 'Date' column to a Date object
MSFT$Date <- as.Date(MSFT$Date)

ggplot(MSFT, aes(x = Date, y = Close)) +
  geom_line(color = "green", size = 1.2) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  labs(
    title = "Microsoft Stock Price Over Time",
    subtitle = "Based on Close Price",
    x = "Date",
    y = "Close Price ($)"
  ) +
  theme_light(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 14),
    axis.title = element_text(face = "bold", size = 14)
  ) +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y") 
```

Microsoft has understood the impact of ChatGPT and they have even partnered with Open AI to bring GPT smarts to their Bing search engine that has been upgraded to be called "Bing Chat". 
Bing chat is Microsoft's big generative AI product and because of it their stock is doing well. Since the release of Bing AI in Feburary 7th 2023, Microsoft's stock has been on a steady 
incline. Finally, we look at Google's stock 

```{r}
GOOG <- read.csv("GOOGL.csv")

# Convert the 'Date' column to a Date object
GOOG$Date <- as.Date(GOOG$Date)

# Create a line plot using ggplot2 with a red line
ggplot(GOOG, aes(x = Date, y = Close)) +
  geom_line(color = "red", size = 1.2) + # Adjusted color and size for better visualization
  labs(
    title = "Google Stock Price Over Time",
    subtitle = "Based on Close Price", 
    x = "Date",
    y = "Close Price ($)"
  ) +
  theme_light(base_size = 12) + # Switched to theme_light for a cleaner look
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 14),
    axis.title = element_text(face = "bold", size = 14),
  )
```

ChatGPT took the world by storm even existing giants such as Google were blindsided. Their stock begins with a decline since November when ChatGPT was released. Since people understood
why would you want to go through the search results yourself on Google and watch advertisements if ChatGPT could just give you the answer. Then in February, Google released their answer
to ChatGPT, Bard. Bard is Google's generative AI that is able to search the web. However, Bard got off to a rocky start since it got a fact wrong in Google's promotional material for it.
Unfortunately, the data here is monthly otherwise during the month of February, Google's stock was in turmoil. However, since then Google's stock has stabilized and seems to be on a steady incline. 

# Results

The analysis carried out on the diverse datasets presents a multi-faceted examination of the influence AI tools, particularly ChatGPT, have wielded on the educational and technological sectors.
The examination spanned through a gamut of areas including students' perceptions on AI, stock market dynamics of education-centric firms like Chegg, and the proliferation of ChatGPT users post 
its inception. Initially, the student survey illustrated a substantial consensus among participants regarding the potential impact of AI on the educational sector. A significant portion of 
respondents (approximately 2/3) anticipated a notable influence, reflecting a general optimism towards AI's role in education. Further, a high frequency of students rating the usefulness of AI 
in education at a perfect score of 10, underlines a strong belief in AI's potential to augment educational processes. On the corporate front, the case of Chegg provided a useful lens through 
which to observe market reactions to the advent of ChatGPT. The analysis revealed a clear inverse correlation between ChatGPT's user growth and Chegg's stock price post-November 2022. The 
graphical representation depicted a pronounced decline in Chegg's stock price concomitant with a steep ascend in ChatGPT's user base, marking a pivotal shift in the educational help domain.
Moreover, the rapidity with which ChatGPT amassed a user base of one million in a mere five days, juxtaposed against other popular services, accentuates its meteoric rise and the consequential 
ripple effect through the tech and education sectors. Furthermore, the stock performance of companies such as Nvidia, Microsoft, and Google, each at different junctures in their AI journey, 
underscored the financial ramifications of AI adoption. Nvidia, having embraced the AI wave early, saw a remarkable stock performance. Conversely, Google experienced a temporary stock 
decline post ChatGPT's release, which saw rectification upon the launch of its own AI tool, Bard.

# Conclusions

The comprehensive analysis encapsulated in this project underscores the pivotal and transformative role AI tools, epitomized by ChatGPT, have played in redefining the landscapes of education and 
technology sectors. The stark contrast in market performance between companies that have integrated AI into their operational or product frameworks and those that haven't, elucidates a clear 
competitive advantage accrued by early adopters of AI technology. The apparent synergy between AI adoption and enhanced market performance, as seen in the cases of Nvidia and Microsoft post their 
AI ventures, delineates a potential roadmap for other entities in the tech and education sectors aiming for robust market performance and enhanced service delivery. Moreover, the student-centric 
perspective gleaned from the survey data, paints a hopeful picture of AI's role in augmenting educational processes, aligning with the market trends observed. The decline in stock value for 
traditional educational help services like Chegg, juxtaposed against the ascension of AI-powered platforms like ChatGPT, underscores a paradigm shift favoring AI-driven solutions. Conclusively, the 
transformative potential of AI is not only vast but seems to be a critical driver for success in the contemporary digital ecosystem. The findings beckon a closer integration of AI tools across sectors, 
to not only remain competitive but to drive innovation and improve service delivery in this burgeoning digital age.

# Further Study

AI is a quickly evolving subject and there is still much that needs to be studied about it. A broader understanding of AI's footprint could be garnered through a comparative analysis across sectors, 
delving into AI's role in healthcare, agriculture, public administration, among others, juxtaposed against its impact in education. Moreover, the ethical implications of AI deployment, especially in 
education, beckon rigorous scrutiny; issues surrounding data privacy, security, and inherent biases in AI responses could have far-reaching consequences. Concurrently, the policy and regulatory impact 
on AI's integration in various sectors poses a significant area of exploration; evolving policies could either foster or stifle innovation, and understanding this dynamic is crucial. Furthermore, the 
economic repercussions of AI adoption cannot be overstated. A meticulous examination of job displacement, market dynamics, financial feasibility for companies and end-users, alongside AI's broader 
economic implications, could provide invaluable insights. These multifaceted studies could elucidate the ripple effects of AI across the societal and economic spectrum, laying the groundwork for 
informed decision-making and strategic planning in harnessing AI's potential responsibly. What an exciting time to be alive!

# Appendix

You can find the datasets used in this project in the following sites

- [Students Survey on AI in Education](https://www.kaggle.com/datasets/gianinamariapetrascu/survey-on-students-perceptions-of-ai-in-education)
- [Chegg Stock Price](https://finance.yahoo.com/quote/CHGG/history?p=CHGG)
- [ChatGPT User Growth](https://nerdynav.com/chatgpt-statistics/)
- [NVIDIA Stock Price](https://finance.yahoo.com/quote/NVDA/history?p=NVDA)
- [Microsoft Stock Price](https://finance.yahoo.com/quote/MSFT/history?p=MSFT)
- [Google Stock Price](https://finance.yahoo.com/quote/GOOGL/history?p=GOOGL)

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```